---
title: "R Notebook"
output: html_notebook
---

```{r}
library (tidyverse)
library(here)
```

```{r}
candy_2015 <- read_csv(here::here('clean_data/clean_data_2015'))
candy_2016 <- read_csv(here::here('clean_data/clean_data_2016'))
candy_2017 <- read_csv(here::here('clean_data/clean_data_2017'))
```

```{r}
test_join_2015_2016 <- full_join(candy_2015, candy_2016)
test_join_2015_2016_2017 <- full_join(test_join_2015_2016, candy_2017)
full_candy <- test_join_2015_2016_2017
write.csv(full_candy, here::here("clean_data/full_candy"), row.names = FALSE)


```

```{r}
full_candy %>% 
  names()
```


Q1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
 
  
candies_only <- select(full_candy, - timestamp, -age, -going_out, -year,
                       -hugs_actual_physical_hugs, -state_province_county_etc, -country, -gender)
names <- names(candies_only)

 total <-sum(!is.na(candies_only))

total_number_of_candy_rating <-  paste('the total number of candy rating is : ',  total , sep = '')
                                        
total_number_of_candy_rating
```

Q2. What was the average age of people who are going out trick or treating?

```{r}

going_out <- subset(full_candy, going_out == "Yes" & age >= 1 & age <= 100)

 mean_age <- mean(going_out$age, na.rm = TRUE)

avg_age_of_people_going_out <- paste('the average of people going out trick or treating is ', 
                               mean_age, sep = '' )

avg_age_of_people_going_out
```

Q3. What was the average age of people who are not going trick or treating?
```{r}
going_out_no <- subset(full_candy, going_out == "No" & age >= 1 & age <= 100)

mean_age_1 <- mean(going_out_no$age, na.rm = TRUE)

avg_age_of_people_not_going_out <- paste('the average of people not going out trick or treating is ', 
                               mean_age_1, sep = '' )

avg_age_of_people_not_going_out
```
Q4. For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candies_only

# pivot longer on the candies columns

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:98),
                                   names_to = "candies",
                                   values_to = "rating")

# counting number of rating per candy
 
rating_count <- candies_only_long %>% 
  count(candies, rating)  


joy <- rating_count %>% 
  slice_max(rating == "JOY") %>% 
  arrange(desc(n))

meh <- rating_count %>% 
  slice_max(rating == "MEH") %>% 
  arrange(desc(n))

despair <- rating_count %>% 
  slice_max(rating == "DESPAIR") %>% 
  arrange(desc(n))

joy
meh
despair

answer <- paste( 'Joy: any_full_sized_candy_bar, Meh: lollipops, Despair: broken_glow_stick ')
                   
                   
answer
 
```


Q5.How many people rated Starburst as despair?

```{r}
rating_count %>% 
  filter(candies == "starburst",rating == "DESPAIR") 

 
```

 For the next three questions, count despair as -1, joy as +1, and meh as 0.

Q6. What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
## reassigning and re positioning year, country and gender to the data set

full_candy_new <- select(full_candy, - timestamp, -age, -going_out, 
                       -hugs_actual_physical_hugs, -state_province_county_etc) %>% 
  mutate(year = as.character(year)) %>% 
  relocate( year, .after = take_5) %>% 
  relocate(gender, .before = year) %>% 
  relocate(country, .after = year)

# pivoting the full_candy_new table 

full_candy_new_long  <- pivot_longer(full_candy_new,
                                   cols = c(1:98),
                                   names_to = "candies",
                                   values_to = "rating")

# counting the rating and converting gender to NA where the answers were not wrong

rating_count_full_candy <- full_candy_new_long %>% 
  count(candies, rating, year, gender,country)  %>% 
  mutate ( gender = case_when(gender == "I'd rather not say" ~ NA_character_ ,
                              gender == "Other" ~ NA_character_,
                              gender == "NA" ~ NA_character_,
                              TRUE ~ gender))

# converting the rating column to -1. 1, and 0 

rating_count_full_candy_new <- rating_count_full_candy %>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

 # converting rating column to numeric

rating_count_full_candy_new_numeric <- rating_count_full_candy_new %>% 
  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy per gender

popular_candy_by_gender <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, gender) %>%
summarise(sum_rating = sum(new_rating)) 
  

# filtering by gender and reordering the table descendant to see the highest rated candy

male_popular_candy <- popular_candy_by_gender %>% 
 filter( gender == "Male") %>% 
arrange(desc(sum_rating))

female_popular_candy <- popular_candy_by_gender %>% 
 filter( gender == "Female") %>% 
arrange(desc(sum_rating))

male_popular_candy
female_popular_candy

```

Q7. What was the most popular candy bar in each year?

```{r}
popular_candy_by_year <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, year) %>%
summarise(sum_rating = sum(new_rating)) 


# filtering by year and reordering the table descendant to see the highest rated candy

popular_candy_2015 <- popular_candy_by_year %>% 
 filter(year == "2015") %>% 
arrange(desc(sum_rating))

popular_candy_2016 <- popular_candy_by_year %>% 
 filter(year == "2016") %>% 
arrange(desc(sum_rating))

popular_candy_2017 <- popular_candy_by_year %>% 
 filter(year == "2017") %>% 
arrange(desc(sum_rating))

popular_candy_2015
popular_candy_2016
popular_candy_2017

```
Q8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

```{r}
popular_candy_by_country <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, country) %>%
summarise(sum_rating = sum(new_rating)) 

popular_candy_USA <- popular_candy_by_country %>% 
 filter(country == "USA") %>% 
arrange(desc(sum_rating))

popular_candy_Canada <- popular_candy_by_country %>% 
 filter(country == "Canada") %>% 
arrange(desc(sum_rating))


popular_candy_UK <- popular_candy_by_country %>% 
 filter(country == "UK") %>% 
arrange(desc(sum_rating))


popular_candy_USA
popular_candy_Canada
popular_candy_UK
```
 
 
 Q8. all other countries
```{r}

popular_candy_by_country_other <- rating_count_full_candy_new_numeric %>% 
  filter(!is.na(country)) %>%
  filter(country != "USA" & country != "Canada" & country != "UK") %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies) %>%
 summarise(sum_rating = sum(new_rating)) %>% 
arrange(desc(sum_rating))  
  
  


 all_other <- max(popular_candy_by_country_other$sum_rating) 
  
 question_8_for_all_other <- paste("The most popular candy for all other country is any_full_sized_candy_bar with a rating of ",  all_other, sep = ''  )
 
  question_8_for_all_other
```

