---
title: "R Notebook"
output: html_notebook
---

```{r}
library (tidyverse)
library(here)
```

```{r}
candy_2015 <- read_csv('clean_data/clean_data_2015')
candy_2016 <- read_csv('clean_data/clean_data_2016')
candy_2017 <- read_csv('clean_data/clean_data_2017')

```

```{r}
test_join_2015_2016 <- full_join(clean_data_2015, clean_data_2016)
test_join_2015_2016_2017 <- full_join(test_join_2015_2016, clean_data_2017)
full_candy <- test_join_2015_2016_2017

```



Q1. What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)

```{r}
full_candy %>% 
  names()

candies_only <- select(full_candy, - timestamp, -age, -going_out, -year,
                       -hugs_actual_physical_hugs, -state_province_county_etc, -country, -gender)
names(candies_only)

sum(!is.na(candies_only))

```

```{r}
joy <- sum(str_count(full_candy, "JOY")) # 277,361
 meh <- sum(str_count(full_candy, "MEH")) # 79,418
despair <- sum(str_count(full_candy, "DESPAIR")) # 274,993

total = joy + meh + despair
```

Q2. What was the average age of people who are going out trick or treating?
```{r}

going_out <- subset(full_candy, going_out == "Yes" & age >= 1 & age <= 100)

mean(going_out$age, na.rm = TRUE)


```
Q3. What was the average age of people who are not going trick or treating?
```{r}
going_out_no <- subset(full_candy, going_out == "No" & age >= 1 & age <= 100)

mean(going_out_no$age, na.rm = TRUE)

```
Q4. For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r}
candies_only

# pivot longer on the candies columns

candies_only_long  <- pivot_longer(candies_only,
                                   cols = c(1:98),
                                   names_to = "candies",
                                   values_to = "rating")

# counting number of rating per candy
 
rating_count <- candies_only_long %>% 
  count(candies, rating)  


joy <- rating_count %>% 
  slice_max(rating == "JOY") %>% 
  arrange(desc(n))

meh <- rating_count %>% 
  slice_max(rating == "MEH") %>% 
  arrange(desc(n))

despair <- rating_count %>% 
  slice_max(rating == "DESPAIR") %>% 
  arrange(desc(n))

joy
meh
despair
all

 all <- rating_count %>% 
  slice_max(rating ) %>% 
  arrange(desc(n))

```


Q5.How many people rated Starburst as despair?

```{r}
rating_count %>% 
  filter(candies == "starburst",rating == "DESPAIR") 
  
```


 For the next three questions, count despair as -1, joy as +1, and meh as 0.

Q6. What was the most popular candy bar by this rating system for each gender in the dataset ?

```{r}
# changing the rating column to 1, -1 and 0

rating_count_new <- rating_count %>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

 # converting rating column to numeric

rating_count_new_numeric <- rating_count_new %>% 
  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy

rating_count_new_numeric_sum <- rating_count_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies) %>% 
summarise(sum_rating = sum(new_rating))

# reordering the table descendant to see the highest rated candy

rating_count_new_numeric_sum_ordered <- rating_count_new_numeric_sum %>% 
  arrange(desc(sum_rating))

rating_count
rating_count_new
rating_count_new_numeric
rating_count_new_numeric_sum
rating_count_new_numeric_sum_ordered
```

Q7. What was the most popular candy bar in each year?

```{r}
full_candy_new <- select(full_candy, - timestamp, -age, -going_out, 
                       -hugs_actual_physical_hugs, -state_province_county_etc) %>% 
  mutate(year = as.character(year)) %>% 
  relocate( year, .after = take_5) %>% 
  relocate(gender, .before = year) %>% 
  relocate(country, .after = year)


full_candy_new_long  <- pivot_longer(full_candy_new,
                                   cols = c(1:98),
                                   names_to = "candies",
                                   values_to = "rating")

#######################################################################
rating_count_full_candy <- full_candy_new_long %>% 
  count(candies, rating, year, gender,country)  %>% 
  mutate ( gender = case_when(gender == "I'd rather not say" ~ NA_character_ ,
                              gender == "Other" ~ NA_character_,
                              gender == "NA" ~ NA_character_,
                              TRUE ~ gender))

rating_count_full_candy_new <- rating_count_full_candy %>% 
  mutate(rating = case_when(rating == "DESPAIR" ~ "-1",
                            rating == "JOY" ~ "1",
                            rating == "MEH" ~ "0",
                            TRUE ~ rating))

 # converting rating column to numeric

rating_count_full_candy_new_numeric <- rating_count_full_candy_new %>% 
  mutate(rating = as.numeric(rating))

# calculating the sum of rating for each candy

rating_count_full_candy_new_sum <- rating_count_full_candy_new_numeric %>% 
  filter( rating == -1 | rating == 1 |rating == 0) %>% 
  mutate( new_rating = rating * n ) %>% 
group_by(candies, gender) %>%
summarise(sum_rating = sum(new_rating)) 
  

# reordering the table descendant to see the highest rated candy

rating_count_full_candy_new_sum %>% 
 #slice_max( gender == "Male") %>% 
  filter(gender == "Female") %>% 
arrange(desc(sum_rating))

rating_count
rating_count_new
rating_count_new_numeric
rating_count_new_numeric_sum
rating_count_new_numeric_sum_ordered


```


```{r}
u <- unique(full_candy_new$gender)
```


